---
title: "March 01-07"
author: "Yumou"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Factor variables

what are the types of variables in the "fbi" data?

```{r, message = FALSE}
library(classdata)
library(tidyverse)

str(fbi)
fbi1 = fbi
```

Make "year", "state" and "type" as factor variables:

```{r}
fbi1$state = as.factor(fbi1$state)
fbi1$type.factor =  as.factor(fbi1$type)
fbi1$state_abbr = as.factor(fbi1$state_abbr)
fbi1$year.factor = as.factor(fbi1$year)
str(fbi1)
```

Boxplots for factor variables:

```{r, warning = FALSE, message = FALSE}
twoyear = dplyr::filter(fbi1, year %in% c(1981, 2020))

ggplot(twoyear, aes(x = year, y = count)) + geom_boxplot()

ggplot(twoyear, aes(x = year.factor, y = count)) + geom_boxplot()
```

Transform factor to numeric variable

```{r}
b1 = as.numeric(fbi1$year.factor) # not working

b2 = as.numeric(as.character(fbi1$year.factor)) # works
```

## levels of factors

Manually reorder the levels of a factor (two ways):

```{r}
levels(fbi1$type.factor)
ggplot(fbi1, aes(x = type.factor)) + geom_bar(aes(weight = count)) + coord_flip() # not ordered

fbi1$type.reorder = factor(fbi1$type, 
                  levels = c("larceny", "burglary", 
                     "motor_vehicle_theft", "aggravated_assault", 
                     "robbery", "arson", "rape_legacy", "rape_revised", "homicide")
                  )
levels(fbi1$type.reorder)

ggplot(fbi1, aes(x = type.reorder)) + geom_bar(aes(weight = count)) + coord_flip() # works
```

The following code would not work:

```{r}
fbi1$type.reorder.2 = fbi1$type.factor
levels(fbi1$type.reorder.2)
levels(fbi1$type.reorder.2) = c("LA", "BU", 
                     "MVT", "AS", 
                     "RB", "AR", "RL", "RV", "HO")
levels(fbi1$type.reorder.2)

ggplot(fbi1, aes(x = type.reorder.2)) + geom_bar(aes(weight = count)) + coord_flip() # not working
```

"reorder()" function:

```{r}
levels(reorder(fbi1$type, fbi1$count, na.rm = TRUE))

A1 = dplyr::filter(fbi, type == "aggravated_assault")
mean(A1$count, na.rm = TRUE)

A2 = dplyr::filter(fbi, type == "arson")
mean(A2$count, na.rm = TRUE)

A3 = dplyr::filter(fbi, type == "robbery")
mean(A3$count, na.rm = TRUE)

# reorder based on the summation of "count" variable
levels(reorder(fbi1$type, fbi1$count, FUN = sum, na.rm = TRUE))
```

## Practice for factor variable

1. Introduce a rate of the number of reported offenses by population into the fbi data. You could use the Ames standard to make values comparable to a city of the size of Ames (population ~70,000).

```{r}
fbi1$rate = fbi1$count / fbi1$population * 70000
```


2. Plot boxplots of crime rates by different types of crime. How can you make axis text legible?

```{r, warning=FALSE}
ggplot(fbi1, aes(x = type, y = rate)) + geom_boxplot() + coord_flip()
```


3. Reorder the boxplots of crime rates, such that the boxplots are ordered by their medians.

```{r, warning=FALSE}
fbi1$type.reorder.median = reorder(fbi1$type, fbi1$count, FUN = median, na.rm = TRUE)

ggplot(fbi1, aes(x = type.reorder.median, y = rate)) + geom_boxplot() + coord_flip()
```

4. For one type of crime (subset or filter) plot boxplots of rates by state, reorder boxplots by median crime rates

```{r}
fbi2 = dplyr::filter(fbi1, type == "homicide")

ggplot(fbi2, aes(x = state, y = rate)) + geom_boxplot() + coord_flip()

fbi2$homicide.reorder.median = reorder(fbi2$state, fbi2$rate, FUN = median, na.rm = TRUE)

ggplot(fbi2, aes(x = homicide.reorder.median, y = rate)) + geom_boxplot() + coord_flip()

# change "District of Columbia" to "DC"
levels(fbi2$homicide.reorder.median)[52] = "DC"
ggplot(fbi2, aes(x = homicide.reorder.median, y = rate)) + geom_boxplot() + coord_flip()
```

