---
title: "Practice Feb 13-15"
author: "Yumou"
date: "2023-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary statistics of *cities* data

The variable "burglary"
```{r}
library(classdata)
A = cities$Burglary

mean(A, na.rm	= TRUE)
sd(A, na.rm = TRUE)
```


## Questions for *cities* data

Please see the questions in the folder "Questions for cities data" in this repository.

## Making plots for *fbiwide* data

```{r}
library(ggplot2)

ggplot(fbiwide, aes(x = log(burglary), y = log(robbery))) + geom_point()

ggplot(fbiwide, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point()

# Add color by years

ggplot(fbiwide, aes(x = log(burglary), y = log(motor_vehicle_theft), colour = year)) + geom_point()

```

## Questions in the lecture "02_r-graphics"

1. Draw a scatterplot of the log transformed number of burglaries by motor vehicle thefts. Map the state variable to colour. Why is this a terrible idea?

```{r}
ggplot(data = fbiwide, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point(aes(color = state))
```

This is a bad idea because too many states and colors can not separate each of them.

2. Draw a scatterplot of the log transformed number of burglaries by motor vehicle thefts. Map population to size. How do we interpret the output?

```{r}
ggplot(data = fbiwide, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point(aes(color = state, size = population))
```

3. Have a look at the RStudio cheat sheet on visualization. Draw a histogram of the state populations

```{r}
ggplot(data = fbiwide, aes(x = population)) + geom_histogram()
```
The above histogram does not make sense as it includes multiple years for the same state. 

We make the histogram of popualtion for 2019.

```{r}
ggplot(data = fbiwide[fbiwide$year == 2019, ], aes(x = population)) + geom_histogram()
```

4. Compare the log transformed number of burglaries by motor vehicle thefts over years. How to make a nice plot?

```{r}
fbiwide1 = fbiwide
fbiwide1 = fbiwide1[fbiwide1$year >= 1980, ]
fbiwide1 = fbiwide1[fbiwide1$year < 2020, ]
fbiwide1$decade = floor(fbiwide1$year / 10)
fbiwide1$decade = paste0(fbiwide1$decade, "0s")

ggplot(data = fbiwide1, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point(aes(color = decade))
```


5. Compare the log transformed number of burglaries by motor vehicle thefts over States, coloured by years.

```{r}
ggplot(data = fbiwide1, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point(aes(color = decade)) + facet_wrap(.~ state_abbr)
```


6. Now, only focus on comparing California, Colorado, Iowa, Illinois, District of Columbia and New York.

```{r}
state.subset = c("California", "Colorado", "Iowa", "Illinois", "District of Columbia", "New York")

fbiwide2 = fbiwide1[fbiwide1$state %in% state.subset, ]
ggplot(data = fbiwide2, aes(x = log(burglary), y = log(motor_vehicle_theft))) + geom_point(aes(color = decade)) + facet_wrap(.~ state)
```

7. We all know population is an important factor. How to compare different states by standardized population?

```{r}
ggplot(data = fbiwide2, aes(x = log(burglary / population * 66424), y = log(motor_vehicle_theft / population * 66424))) + geom_point(aes(color = decade)) + facet_wrap(.~ state)
```

