---
title: "March 08-22"
author: "Yumou"
date: "2023-03-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 'dplyr' package

### pipe operator '%>%'

Let $f(x) = x^2 + 1$, and $g(x, y) = x + y^2$. Use the pipe oprator to calculate $g(f(1), 2)$.

```{r, message = FALSE}
library(dplyr)

f1 = function(x){
  return(x^2 + 1)
}

g1 = function(x, y){
  return(x + y^2)
}

g1(f1(1), 2)

1 %>% f1() %>% g1(2)
```

How about $g(3, f(1))$? This can not be computed using the pipe operator

```{r}
g1(3, f1(1))

1 %>% f1() %>% g1(3)
```

Let $h(x) = x^3$. Calculate $h(f(1))$?

```{r}
h1 = function(x){
  return(x^3)
}

1 %>% f1() %>% h1()
```

## Practice questions

1. Use the `fbi` data, select the recordings for Iowa, calculate the crime rate, ordered by year, crime rate. Lastly, output the data with year, type of crime and crime rate.

```{r}
library(classdata)

fbi.Iowa = 
fbi %>% filter(state == "Iowa") %>% 
  mutate(crime.rate = count / population * 70000) %>%
  arrange(desc(year), desc(crime.rate)) %>%
  select(year, type, crime.rate)
```

2. Compare the crime rate for Iowa, Nebraska and Minnesota. Please calculate the average, 25% and 75% quantile of crime rate for each type of crime.

```{r, message = FALSE}
fbi.summary = 
fbi %>% filter(state %in% c("Iowa", "Nebraska", "Minnesota")) %>% 
  mutate(crime.rate = count / population * 70000) %>%
  group_by(state, type) %>%
  summarise(average = mean(crime.rate, na.rm = TRUE),
            quan25 = quantile(crime.rate, 0.25, na.rm = TRUE),
            quan75 = quantile(crime.rate, 0.75, na.rm = TRUE))
```

3. Make a bar plot to view the data, facet by the summary statistics.

```{r}
library(ggplot2)

fbi.summary %>% ggplot(aes(x = type, fill = state)) + 
  geom_bar(aes(weight = average), position = "dodge") + coord_flip()

library(reshape2)
fbi.summary %>% melt() %>% 
  ggplot(aes(x = type, fill = state)) + 
  geom_bar(aes(weight = value), position = "dodge") + 
  facet_wrap(~variable) + coord_flip()
```

4. Use mutate to reorder types of crimes by (median) rate. Plot crime rates by type in side-by-side boxplots. Medians of the boxplots should be ordered.

```{r, message=FALSE, warning=FALSE}
fbi %>% mutate(crime.rate = count / population * 70000) %>%
  mutate(type.reorder = factor(type, levels(reorder(type, crime.rate, median, na.rm = TRUE)))) %>%
  ggplot(aes(x = type.reorder, y = crime.rate)) + geom_boxplot() + coord_flip()
```

