---
title: "In-class practice"
author: "Yumou"
date: "2023-04-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time series data

```{r, message = FALSE, warning = FALSE}
library(tidyverse)

data(nasa, package="GGally")
```

Choose two locations in `x` and two locations in `y`

```{r}
nasa %>% filter(x %in% c(1, 2), y %in% c(1, 10)) %>% 
  mutate(y = factor(y), x = factor(x)) %>%
  ggplot(aes(x = time, y = temperature, group = id, color = y)) + 
  geom_line(aes(linetype = x))

```

1. For one location, draw a time line of Ozone over the time frame (time).

```{r}
nasa %>% filter(x == 1, y == 1) %>%
  ggplot(aes(x = date, y = ozone)) + geom_line()
```

2. Plot separate lines for each of the years, i.e. put month on the x-axis and ozone on the y-axis for the same location. Is there a seasonal pattern apparent?

```{r}
library(lubridate)

nasa %>% filter(x == 1, y == 1) %>% 
  ggplot(aes(x = month(date, label = TRUE), y = ozone, group = year, color = factor(year))) + geom_line()
```

3. Pick a location with x in 1:10 and y in 7:10. Plot temperature over time.

```{r}
nasa %>% filter(x %in% c(1 : 10), y %in% c(7 : 10)) %>% 
  mutate(y = factor(y), x = factor(x)) %>%
  ggplot(aes(x = time, y = temperature, group = id, color = x)) + 
  geom_line() + facet_wrap(~y)
```

4. Use `gather()` function to put different variables versus time

```{r}
nasa %>% gather(key = variable, value = value, 7 : 13) %>%
  filter(x == 2, y == 1) %>% 
  ggplot(aes(x = month(date, label = TRUE), y = value, group = year, color = year)) + 
  geom_line() + facet_wrap(~variable, scales = "free_y")
```

## `Box` data

1. For each movie and distributor, find the (1) highest total gross, (2) the last date (and week) the movie was shown in theaters, (3) the gross the movie made in the first week it was released.

```{r}
data(box, package="classdata")

box_summary = box %>% group_by(Movie, Distributor) %>%
  summarise(Gross.final.1 = max(Total.Gross),
            Gross.final.2 = Total.Gross[which.max(Week)],
            Gross.initial.1 = min(Total.Gross),
            Gross.initial.2 = Total.Gross[which.min(Week)],
            Max.Thtrs = max(Thtrs.),
            date.last = max(Date),
            week.count = max(Week)) %>%
  mutate(Gross.final.diff = Gross.final.1 - Gross.final.2,
         Gross.initial.diff = Gross.initial.1 - Gross.initial.2)

box_summary1 = box_summary %>% filter(Gross.final.diff == 0,
                                      Gross.initial.diff == 0,
                                      Max.Thtrs > 100,
                                      week.count < 30)

name.clean = box_summary1$Movie

box1 = box %>% filter(Movie %in% name.clean)
```

```{r, warning = FALSE}
box_summary2 = box_summary1 %>% arrange(desc(Gross.final.1))
box_summary2 = box_summary2[1 : 3, ]

box1 %>% ggplot(aes(x = Date, y = Total.Gross, 
                    group = interaction(Movie, Distributor))) + geom_line() +
  geom_text(aes(x = date.last - 500, y = Gross.final.1, label = Movie, color = Movie), 
            data = box_summary2)
```

```{r, warning = FALSE}
box1 %>% ggplot(aes(x = Date, y = Total.Gross, 
                    group = interaction(Movie, Distributor))) + geom_line() +
  geom_text(aes(x = date.last, y = Gross.final.1, label = Movie), 
            data = box_summary1 %>% filter(Gross.final.1 >= 368049635) )
```

Non-overlapping text:

```{r, warning = FALSE}
box1 %>% ggplot(aes(x = Date, y = Total.Gross, 
                    group = interaction(Movie, Distributor))) + geom_line() +
  ggrepel::geom_text_repel(aes(x = date.last, y = Gross.final.1, label = Movie), 
            data = box_summary1 %>% filter(Gross.final.1 >= 368049635),
            colour="grey50")
```

## Layers

```{r, message = FALSE}
nasa.1 = nasa %>% filter(id == "1-10") %>% 
  filter(temperature == max(temperature))

nasa %>% filter(id == "1-1") %>% 
  ggplot(aes(x = time, y= temperature)) + geom_point() +
  geom_smooth(method="loess", span = 0.25) + 
  geom_point(data = nasa %>% filter(id == "1-10"), 
             aes(x = time, y= temperature), 
             shape = 2, col = 2) + 
  geom_smooth(data = nasa %>% filter(id == "1-10"), 
             aes(x = time, y= temperature), 
             method="loess", span = 0.25, col = 2) + 
  geom_text(data = nasa.1, aes(x = time, y = temperature + 0.5, label = temperature), col = "grey50", size = 2.5)
```

```{r, message = FALSE}
nasa.2 = nasa %>% filter(id == "1-1") %>% 
  mutate(season = "winter") %>%
  mutate(season = replace(season, month %in% c(6, 7, 8), "summer")) %>%
  mutate(season = replace(season, month %in% c(3, 4, 5), "spring")) %>%
  mutate(season = replace(season, month %in% c(9, 10, 11), "fall"))

nasa %>% filter(id == "1-1") %>%
  ggplot(aes(x = time, y= temperature)) + geom_point() +
  geom_smooth(method="loess", span = 0.25) +
  geom_text(aes(x = time, y = temperature + 0.2, label = temperature, colour = season), 
            data = nasa.2, size = 2.5)
```

```{r, message = FALSE}
nasa %>% filter(id %in% c("1-1", "1-10")) %>%
  ggplot(aes(x = time, y= temperature, group = id)) + geom_point(aes(colour=id)) +
  geom_smooth(method="loess", span = 0.25) +
  geom_text(aes(x = time, y= temperature, label=id), 
            data = nasa %>% filter(id %in% c("1-1", "1-10"), time == 50))
```

## Map

```{r}
library(maps)

states = map_data("state")

states %>% ggplot(aes(x = long, y = lat)) + geom_point()

states %>% ggplot(aes(x = long, y = lat)) + geom_path(aes(group = group))

```

