---
title: "Feb 6 and 8 in-class practice"
author: "Yumou"
date: "2023-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Write a function to compute the Euclidean norm of a vector (square root of sum of squared elements in a vector).
Namely, for a vector $x = (x_1, x_2, \ldots, x_p)$, the Euclidean norm of $x$ is
$$|x|_2 = (x_1^2 + x_2^2 + \ldots + x_p^2)^{1/2}.$$
```{r}
EuclideanNorm = function(yqiu){
  res = sqrt(sum(yqiu^2))
  return(res)
}

EuclideanNorm(c(1 : 5))

EuclideanNorm(-10)
```

## Question 2

Write a loop function to calculate the Euclidean norm of a vector

```{r}
x = c(3 : 16)

EuclideanNorm(x)

d = 0
for (i in 1 : length(x)){
  d = d + x[i]^2
  cat("Iteration = ", i, ", value of x =", x[i], "value of d = ", d, "\n")
}
sqrt(d)

```

## Question 3

Use loop to write the function for Euclidean norm

```{r}
EuclideanNormLoop = function(x){
  d = 0
  for (i in 1 : length(x)){
    d = d + x[i]^2
  }
  return(sqrt(d))
}

EuclideanNormLoop(c(-10 : 10))

EuclideanNorm(c(-10 : 10))
```

## Question 4

Write a function to calculate sample variance of a vector $x$

```{r}
myvar = function(x){
  n = length(x)   # sample size of x
  xbar = sum(x) / n
  x.centered = x - xbar
  res = sum(x.centered^2) / (n - 1)
  return(res)
}

myvar(c(-5 : 20))

var(c(-5 : 20)) # R build-in variance function

```

## Question 5 (simulation example)

Suppose we have data from a Normal distribution with mean $\mu = 5$ and variance $\sigma^2 = 1$. We want to estimate $\theta = \mu^2$ (true value is 25).

Suppose $X_1, \ldots, X_n$ are the data. We have two estimators:

1. $\hat{\theta}_1 = \bar{X}^2$ where $\bar{X} = \sum_{i = 1}^{n} X_i$

2. $$\hat{\theta}_2 = \frac{1}{n(n-1)}\sum_{i \neq j} X_i X_j = \frac{1}{n(n-1)} \bigg\{ \bigg( \sum_{i = 1}^{n} X_i \bigg)^2 - \sum_{i = 1}^{n} X_i^2 \bigg\}$$ 

Which is the estimator better?

```{r}
theta1 = function(x){
  res = mean(x)^2
  return(res)
}

theta2 = function(x){
  n = length(x)
  res = (sum(x)^2 - sum(x^2)) / (n * (n - 1))
  return(res)
}

# Simulation
# Suppose the sample size is 20 (n = 20)
# Simulation from normal distribution with mean 5 and variance 1
# Simualtion 1000 times

diff1 = c()
diff2 = c()

for (rep in 1 : 1000){
  z = rnorm(20, 5, 1)
  diff1[rep] = theta1(z) - 25
  diff2[rep] = theta2(z) - 25
}

hist(diff1)
hist(diff2)

MSE1 = sqrt(mean(diff1^2))
MSE2 = sqrt(mean(diff2^2))

MSE1

MSE2
```



