---
title: "Feb 22-27"
author: "Yumou"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logic operations

For a given vector $a = (1, 15, 3, 20, 5, 8, 9, 10, 1, 3)$,
```{r}
library(classdata)
library(ggplot2)

a = c(1, 15, 3, 20, 5, 8, 9, 10, 1, 3)
```

1. less than 20

```{r}
a < 20
```

2. squared value is at least 100 or less than 10

```{r}
(a^2 >= 100) | (a^2 < 10)
a^2
```


3. equals 1 or 3

```{r}
(a == 1) | (a == 3)
a %in% c(1, 3)
```

4. even

```{r}
a
(a %% 2) == 0
((a / 2) - floor(a / 2)) == 0
```

## "filter" function

```{r, message = FALSE}
library(tidyverse)

A1 = dplyr::filter(fbi, year == 2014, state == "Iowa")
```

1. Get a subset of all crimes in Iowa, Plot incidences/rates for one type of crime over time.

```{r, warning = FALSE}
A2 = dplyr::filter(fbi, state == "Iowa")
ggplot(data = A2, aes(x = year, y = count / population)) + 
  geom_point() + geom_line() + 
  facet_wrap(~type, scales = "free_y")
```

2. Get a subset of all crimes in 2009. Plot the number or rate for one type of crime by state.

```{r}
A3 = dplyr::filter(fbi, year == 2009, type %in% c("homicide", "arson", "robbery"))
ggplot(data = A3, aes(x = state)) + 
  geom_bar(aes(weight = count / population)) + coord_flip() + 
  facet_wrap(~type, scale = "free")
```

3. Get a subset of the data that includes number of homicides for the last five years. Find the rate of homicides, extract all states that have a rate of greater than 90% of the rates across all states, and plot 

```{r}
A4 = dplyr::filter(fbi, year %in% c(2016 : 2020), type == "homicide")
A4$homicide.rate = A4$count / A4$population * 66000
quan09 = quantile(A4$homicide.rate, 0.9)
A5 = dplyr::filter(A4, homicide.rate > quan09)

ggplot(A5, aes(x = state)) + geom_bar(aes(weight = homicide.rate)) + 
  facet_wrap(~year) + coord_flip()
```

## "replace" function

Replace the problematic value in life expectancy for Canada in 1957 in the gapminder data set by 69.96.

```{r}
B = read.csv("gapminder-5060.csv")
dplyr::filter(B, country == "Canada", year == 1957)

B$lifeExp = replace(B$lifeExp, (B$country == "Canada") & (B$year == 1957), 69.96)
dplyr::filter(B, country == "Canada", year == 1957)
```

