---
title: "Feb 22-27"
author: "Yumou"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logic operations

For a given vector $a = (1, 15, 3, 20, 5, 8, 9, 10, 1, 3)$,
```{r}
library(classdata)
library(ggplot2)

a = c(1, 15, 3, 20, 5, 8, 9, 10, 1, 3)
```

1. less than 20

```{r}
a < 20
```

2. squared value is at least 100 or less than 10

```{r}
(a^2 >= 100) | (a^2 < 10)
a^2
```


3. equals 1 or 3

```{r}
(a == 1) | (a == 3)
a %in% c(1, 3)
```

4. even

```{r}
a
(a %% 2) == 0
((a / 2) - floor(a / 2)) == 0
```

## "filter" function

```{r, message = FALSE}
library(tidyverse)

A1 = dplyr::filter(fbi, year == 2014, state == "Iowa")
```

1. Get a subset of all crimes in Iowa, Plot incidences/rates for one type of crime over time.

```{r, warning = FALSE}
A2 = dplyr::filter(fbi, state == "Iowa")
ggplot(data = A2, aes(x = year, y = count / population)) + 
  geom_point() + geom_line() + 
  facet_wrap(~type, scales = "free_y")
```

2. Get a subset of all crimes in 2009. Plot the number or rate for one type of crime by state.

```{r}
A3 = dplyr::filter(fbi, year == 2009, type %in% c("homicide", "arson", "robbery"))
ggplot(data = A3, aes(x = state)) + 
  geom_bar(aes(weight = count / population)) + coord_flip() + 
  facet_wrap(~type, scale = "free")
```

3. Get a subset of the data that includes number of homicides for the last five years. Find the rate of homicides, extract all states that have a rate of greater than 90% of the rates across all states, and plot 

```{r}
A4 = dplyr::filter(fbi, year %in% c(2016 : 2020), type == "homicide")
A4$homicide.rate = A4$count / A4$population * 66000
quan09 = quantile(A4$homicide.rate, 0.9)
A5 = dplyr::filter(A4, homicide.rate > quan09)

ggplot(A5, aes(x = state)) + geom_bar(aes(weight = homicide.rate)) + 
  facet_wrap(~year) + coord_flip()
```

## "replace" function

Replace the problematic value in life expectancy for Canada in 1957 in the gapminder data set by 69.96.

```{r}
B = read.csv("gapminder-5060.csv")
dplyr::filter(B, country == "Canada", year == 1957)

B$lifeExp = replace(B$lifeExp, (B$country == "Canada") & (B$year == 1957), 69.96)
dplyr::filter(B, country == "Canada", year == 1957)
```

## Factor variables

what are the types of variables in the "fbi" data?

```{r}
str(fbi)
fbi1 = fbi
```

Make "year", "state" and "type" as factor variables:

```{r}
fbi1$state = as.factor(fbi1$state)
fbi1$type=  as.factor(fbi1$type)
fbi1$state_abbr = as.factor(fbi1$state_abbr)
fbi1$year.factor = as.factor(fbi1$year)
str(fbi1)
```

Boxplots for factor variables:

```{r, warning = FALSE, message = FALSE}
twoyear = dplyr::filter(fbi1, year %in% c(1981, 2020))

ggplot(twoyear, aes(x = year, y = count)) + geom_boxplot()

ggplot(twoyear, aes(x = year.factor, y = count)) + geom_boxplot()
```

Transform factor to numeric variable

```{r}
b1 = as.numeric(fbi1$year.factor) # not working

b2 = as.numeric(as.character(fbi1$year.factor)) # works
```

## levels of factors

Manually reorder the levels of a factor (two ways):

```{r}
levels(fbi1$type)
ggplot(fbi1, aes(x = type)) + geom_bar(aes(weight = count)) + coord_flip() # not ordered

fbi1$type.reorder = factor(fbi1$type, 
                  levels = c("larceny", "burglary", 
                     "motor_vehicle_theft", "aggravated_assault", 
                     "robbery", "arson", "rape_legacy", "rape_revised", "homicide")
                  )
levels(fbi1$type.reorder)

ggplot(fbi1, aes(x = type.reorder)) + geom_bar(aes(weight = count)) + coord_flip() # works

fbi1$type.reorder.2 = fbi1$type
levels(fbi1$type.reorder.2)
levels(fbi1$type.reorder.2) = c("larceny", "burglary", 
                     "motor_vehicle_theft", "aggravated_assault", 
                     "robbery", "arson", "rape_legacy", "rape_revised", "homicide")
levels(fbi1$type.reorder.2)

ggplot(fbi1, aes(x = type.reorder.2)) + geom_bar(aes(weight = count)) + coord_flip() # not working

```

"reorder()" function:

```{r}
levels(reorder(fbi1$type, fbi1$count, na.rm = TRUE))

A1 = dplyr::filter(fbi, type == "aggravated_assault")
mean(A1$count, na.rm = TRUE)

A2 = dplyr::filter(fbi, type == "arson")
mean(A2$count, na.rm = TRUE)

A3 = dplyr::filter(fbi, type == "robbery")
mean(A3$count, na.rm = TRUE)

# reorder based on the summation of "count" variable
levels(reorder(fbi1$type, fbi1$count, FUN = sum, na.rm = TRUE))
```

## Practice for factor variable

1. Introduce a rate of the number of reported offenses by population into the fbi data. You could use the Ames standard to make values comparable to a city of the size of Ames (population ~70,000).

```{r}
fbi1$rate = fbi1$count / fbi1$population * 70000
```


2. Plot boxplots of crime rates by different types of crime. How can you make axis text legible?

```{r, warning=FALSE}
ggplot(fbi1, aes(x = type, y = rate)) + geom_boxplot() + coord_flip()
```


3. Reorder the boxplots of crime rates, such that the boxplots are ordered by their medians.

```{r, warning=FALSE}
fbi1$type.reorder.median = reorder(fbi1$type, fbi1$count, FUN = median, na.rm = TRUE)

ggplot(fbi1, aes(x = type.reorder.median, y = rate)) + geom_boxplot() + coord_flip()
```

4. For one type of crime (subset or filter) plot boxplots of rates by state, reorder boxplots by median crime rates